---
title: "Class notes 11 14"
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(readxl)

parenting <- read_excel("parenting.xlsx")

```

## Linear Regression

Considering the relationship between two quantitative variables. Regression considers a possible __linear__ relationship. 

for examples is there a relationship between `baby.sleep` and `dan.grump`

$H_o$: There is a no linear relationship between the variables (the slope of the line of best fit is 0)


$H_a$: there is a linear relationship between the variable (the slope of the line fit is non-zero)

A visualization is essential when doing a regression because you have to check if the relationship appears linear. 

```{r}
parenting %>% 
  ggplot(aes(x = baby.sleep, #what is causing an effect
             y = dan.grump)) + #what is being affected 
  geom_point() + 
  geom_smooth(method = 'lm', #linear model
              se = FALSE) # no error ribbon
```
Run the test with the `lm` command. 

```{r}
model <- lm(dan.grump ~ baby.sleep,
   data = parenting)

summary(model)
```

The model is `dan.grump = 85.7817 - 2.7421 * baby.sleep` The p-value is strong evidence against the null. We can conclude that in the larger population (all days after the baby sleeps), the baby's sleep does impact Danielle's grumpiness. 