---
title: "Class_Notes_9_08"
author: "Ilana Berlin"
date: "8 September 2025"
format: html
embed-resources: true
---
Bar charts are used to see 1-2 categorical variables. 

Explore the `carseats` dataset from the `ISLR2` packages. Install the package in the console, __never__ install in the quarto doc. 

```{r}
#| message: false
library(ISLR2)
library(tidyverse)
```

Use the `data()` command to view all available datasets

```{r}
glimpse(Carseats)
```

Lets look at `SheveLoc` and `USA` 

```{r}
ggplot(Carseats, aes(x = ShelveLoc, fill = US)) + 
  geom_bar() + 
  scale_fill_brewer(palette = "Dark2")
```
This plot is good for showing the relative counts of `ShelveLoc`. It also shows the relative proportions of `US` but that is a little more unclear. Switch which is which to change emphasis. 

```{r}
ggplot(Carseats, aes(fill = ShelveLoc, x = US)) + 
  geom_bar() + 
  scale_fill_brewer(palette = "Dark2")
```
Let's stress relative proportions with the same data and the same aesthetics. (Change the geom)
```{r}
ggplot(Carseats, aes(fill = ShelveLoc, x = US)) + 
  geom_bar(position = "fill") + 
  scale_fill_brewer(palette = "Dark2")
```

