---
title: "Class_Notes_09_10"
format: html
author: "Ilana Berlin"
date: "10 September 2025"
embed-resources: true
---
```{r}
#| label: LoadLibs
#| message: false
library(tidyverse)

```

use `#| label` to name chucks and maintain internal organization 

## Faceting

Break down `penguins` bill length and bill depth by species. 

```{r}
#| label: Facet_wrap
#| warning: false
ggplot(penguins, aes(x = bill_len,
                     y = bill_dep)) +
  geom_point() + 
  facet_wrap(~species, 
             ncol = 1)

```

`facet_wrap` create a new subplot for each variable but uses the same range for each axis. Use `nrow` and `ncol` to specify layout.

Improve the visuals:
```{r}
#| label: Facet_map(goodvis)
#| warning: false
ggplot(penguins, aes(x = bill_len,
                     y = bill_dep, 
                     color = species)) +
  geom_point() + 
  facet_wrap(~species, 
             ncol = 1) + 
  labs(x = 'Bill Length (mm)', 
       y = 'Bill Depth (mm)', 
       color = 'Species',
       title = "Penguin Bill Dimensions",
       caption = "Palmer Station Antarctica LTER and K. Gorman (2020)") + 
  scale_color_brewer(palette = 'Dark2') + 
  theme_minimal() + 
  theme(strip.text = element_blank()) # or theme(legend.position = "none") to remove legend

```
Note: There is missing data about 2 penguines 


Always use `scale_color` or `scale_fill` with palettes `viridis` or `Dark2` to make plots more colorblind friendly. 

Compare `sex`. This is not a good plot, only for example. 

```{r}
#| label: facet_sex
#| warning: false
ggplot(penguins, aes(x = bill_len,
                     y = bill_dep, 
                     color = sex)) +
  geom_point() + 
  facet_wrap(~species, ncol = 1) + 
  labs(x = "Bill Depth (mm)", 
       y = "Bill Length", 
       color = 'Sex') + 
  scale_color_brewer(palette = 'Dark2') +
  theme_minimal()


```
Consider if all variables present in the graph are necessary. 
