---
title: "Class Notes 9_22"
format: html
embed-resources: true
author: Ilana Berlin
date: 22 Sept 2025
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
```

### More on `mutate` 

Change the unit of `bill_len` and `bill_dep` to inches. Multiply by 0.03937008. Create new columns - do not override the old values. Also remove capitalizations to make everything consitent

```{r}
#| label: Convert mm to in

penguins_nonmetric <- penguins %>% 
  mutate(bill_len_in = bill_len * 0.03937008, 
         bill_dep_in = bill_dep * 0.03937008, 
         species = tolower(species), 
         island = tolower(island))

```

### Grouped Summaries 

To start, check out `summarize` (`summarise` also works)

```{r}
#| label: Summarize Total

penguins %>% 
  summarise(bill_len_ave = mean(bill_len, 
                 na.rm = TRUE ), 
            bill_len_sd = sd(bill_len,
                             na.rm = TRUE)) 
```
Use `summarize` to get this data for each different species. The most direct way to do this is using the `.by` command. 


```{r}
#| label: Summarize by species

penguins %>% 
  summarise(bill_len_ave = mean(bill_len, 
                 na.rm = TRUE ), 
            bill_len_sd = sd(bill_len,
                             na.rm = TRUE), 
            .by = species) 
```
You can also use this older but less elegant way. 

```{r}
#| label: Summarize by species - old

penguins %>% 
  group_by(species) %>% 
  summarise(bill_len_ave = mean(bill_len, 
                 na.rm = TRUE ), 
            bill_len_sd = sd(bill_len,
                             na.rm = TRUE))
```
