---
title: "Class_notes_9_24"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
```

### More about `group_by` and `summarize`

Look at `ggplot2::diamonds` (package::function) and get some grouped summaries. Lets find how `price` varies by `cut`. Use count  `n()` to see how datapoints there are in each group. (best practice)

```{r}
#| label: Group by Diamonds

diamonds %>% 
  group_by(cut) %>% 
  summarize(avg_price = mean(price), 
            avg_carat = mean(carat), 
            count = n())


```
 ### Great table (`gt`)

Use `gt::gt` to make tables. 
 
```{r}
#| label: Using gt

diamonds %>% 
  group_by(cut) %>% 
  summarize(avg_price = mean(price), 
            avg_carat = mean(carat), 
            count = n()) %>% 
  gt::gt() #only load the gt function for the package

```
 The `gt` package allows for a lot of customization. Use `?gt` to view the help files for ideas. 
 
 
### Counting Observations 

You can use `count()` as shortcut instead of using `n()` for the counts

```{r}
#| label: Count

diamonds %>% 
  count(cut) %>% 
  gt::gt()
```



 

